---
title: "Data Processing"
output: html_document
editor_options: 
  chunk_output_type: console
---
#Load Packages
```{r}
library(Seurat)
library(harmony)
library(Nebulosa)
library(biomaRt)
library(dplyr)
gene_lists <- readRDS("/premiumDisk/khoh/Github/scTME/data/gene_lists.rds")
source('/premiumDisk/khoh/Github/scTME/R/score_sig.R', echo=TRUE)
```


#@ Lymphocytes subpopulations
```{r}
Lymphocytes<-subset(PDACTrainingMeta,CellType1%in%c("T.Cells","B.Cells","Plasma"))
list_full<-SplitObject(Lymphocytes,split.by = "Dataset")
int.features <- SelectIntegrationFeatures(object.list = list_full, nfeatures = 4000)
Powers <- list_full$Powers
list_full$Powers <- NULL
Lymphocytes <- Seurat::merge(Powers, y = c(list_full),  project = "Lymphocytes", merge.data = TRUE)
VariableFeatures(Lymphocytes) <- int.features
Lymphocytes <- RunPCA(object = Lymphocytes, assay = "SCT", npcs = 30)
Lymphocytes <- RunHarmony(object = Lymphocytes,
                                    assay.use = "SCT",
                                    reduction = "pca",
                                    dims.use = 1:30,
                                    group.by.vars = c("Patient","Phase"),
                                    plot_convergence = TRUE)
ElbowPlot(Lymphocytes,ndims=40,reduction="harmony")
Lymphocytes <- RunUMAP(object = Lymphocytes, assay = "SCT", 
                   reduction = "harmony", 
                   min.dist = 1,
                  negative.sample.rate = 20,
                   n.neighbors = 300,
                   dims = 1:18)
Lymphocytes <- FindNeighbors(object = Lymphocytes, assay = "SCT", reduction = "harmony", dims = 1:18)
Lymphocytes <- FindClusters(object = Lymphocytes, resolution = 1.2)#original
Lymphocytes <- FindClusters(object = Lymphocytes, resolution = 0.5)
DimPlot(Lymphocytes,label=T,group.by = c("CellType2","seurat_clusters"),reduction = "tsne")
Idents(Lymphocytes)<-"seurat_clusters"
Lymphocytes<-Seurat::FindSubCluster(Lymphocytes,cluster=0,graph.name ="SCT_snn",resolution = 0.2)
DimPlot(Lymphocytes,label=T,reduction = "umap",group.by = "sub.cluster")

Lymphocytes<-RunTSNE(Lymphocytes,reduction = "harmony",dims = 1:18)

Lymphocytes<-score_sig(Lymphocytes)
plot_density(Lymphocytes,features =c( "CD8A","CIBERSORT.T.cells.regulatory..Tregs.","CD8B","CIBERSORT.T.cells.CD4.naive","CIBERSORT.T.cells.CD8","Puram_CD8.Tcell","CIBERSORT.NK.cells.activated","NKG7","FOXP3","CD4","Puram_CD4.Tcell","Puram_T.reg","CD79A","Puram_CD6.Texhaust","CD3G","LTB","KLRD1","CD3D","Bernard_Cytotoxic.T"),size = 0.1,reduction = "tsne")

```

#>> Lymphocytes Heatmap
```{r}
Lymphocytes<-BuildClusterTree(Lymphocytes,dims=1:18,reorder = T)
PlotClusterTree(Lymphocytes)
table(Lymphocytes$seurat_clusters)
Idents(Lymphocytes)<-"sub.cluster"

small <- subset(Lymphocytes, downsample = 300)
DefaultAssay(small)<-"RNA"
small<-NormalizeData(small)
small<-ScaleData(small,vars.to.regress = "Dataset",features = rownames(small))
Idents(Lymphocytes)<-"sub.cluster"
fullmarkers <- FindAllMarkers(Lymphocytes,min.pct = 0.2,min.diff.pct = 0.1,test.use = "MAST",only.pos = F)

top10 <- fullmarkers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)
toppval <- fullmarkers %>% group_by(cluster) %>% top_n(n = -5, wt = p_val)
#Idents(small)<-"seurat_clusters"
DoHeatmap(small, features = top10$gene,label=T) + NoLegend()
DoHeatmap(small, features = toppval$gene,label=T) + NoLegend()
DotPlot(Lymphocytes,features = c(unique(top10$gene),"CD4","CD8A","CD8B"),cluster.idents = T)+RotatedAxis()

new.cluster.ids<-c(
"Plasma",#6
"Prolif.Lymph",#7
"Early.CD8",#0_0
"Dysfunctional.CD8",#1
"Progenitor.Lymph",#3
"Cytotoxic.CD8.T",#4
"T.Reg",#5
"Mem.B.Cell",#2
"Naive.CD4.T.Cell",#0_1
"Immature.B.Cell"#8
)

names(new.cluster.ids) <- levels(Lymphocytes)
Lymphocytes <- RenameIdents(Lymphocytes, new.cluster.ids)
Lymphocytes$CellType2<-Idents(Lymphocytes)
#================== Subcluster CD4
Lymphocytes<-FindSubCluster(Lymphocytes,cluster)
#saveRDS(Lymphocytes,"Lymphocytes_Subset2021.rds")

#+=======================================================
